---
title: "N-analysis"
author: "Tuure Parviainen"
date: "January 14, 2017"
output: html_document
---
# Nitrogen analysis

### Abstract
We are looking at the nitrates in soil in a multi-year field growing experiment. The dataset consist of six  different fertilizer treatments. There are four plant treatments. The target variables were nitrate and sulfur in the soil. Leaching of the nitrogen was reduced by added carbon in first year **x** second year **y** and third year **z** amount in autumn. The fertilizers increased sulfphate after fertilization, but in three years there was trace amounts.

Most of the analyses are written in Finnish.

### Legend:

Variables     | Description
----------    | --------------------------------------
ker           | Kerranteet
kas           | Kasvi käsittelyt: "R", "KK", "P", "AK" 
lan           | Lannoite käsittelyt: "S1", "S2", "S3","S4", "Min", "C"
time          | Aika: "29.9.2015", "23.10.2015", "15.11.2015", "22.4.2016", "12.9.2016", "xx.10.2017"
nitN          | Nitraattipitoisuus
TOCN          | Hiilityppisuhde
LisC          | Lisatty kokonaistyppi

## Data exploration

```{r, message=FALSE, warning=FALSE}

require("easypackages") # for loading and installing many packages at one time
packages_to_load <- c("broom", "dplyr", "plyr", "tidyverse", "corrplot", "ggplot2", "GGally", "devtools", "ggthemes")
packages(packages_to_load, prompt = TRUE) # lock'n'load install/load

```

```{r, message=FALSE, warning=FALSE}
Kuitub <- read.table(file = "d1alln2017.txt",
                     header = TRUE,
                     dec = ".")

#Naming the factor levels
#Turning to factors
#Kuitub$lan=factor(Kuitub$lan)
Kuitub$ker=factor(Kuitub$ker)
#Kuitub$kas=factor(Kuitub$kas)
Kuitub$time=factor(Kuitub$time)
#Kuitub$TOCN=factor(Kuitub$TOCN)
#Int to num
Kuitub$Smgkg <- as.numeric(as.character(Kuitub$Smgkg))
Kuitub$KuitLisKokN <- as.numeric(Kuitub$KuitLisKokN)
Kuitub$KuitLisN <- as.numeric(Kuitub$KuitLisN)
Kuitub$TOCN <- as.numeric(Kuitub$TOCN)
Kuitub$TOCNliu <- as.numeric(Kuitub$TOCNliu)
Kuitub$LisC <- as.numeric(Kuitub$LisC)
      
#Giving labes kas
Kuitub$lan <- factor(Kuitub$lan,
                     levels = c(1,2,3,4,5,6),
                     labels = c("S1", "S2", "S3","S4", "Min", "C"))

#Giving labes kas
Kuitub$kas <- factor(Kuitub$kas,
                     levels = c(1,2,3,4),
                     labels = c("R", "KK", "P","AK"))
#Giving labes time
Kuitub$time <- factor(Kuitub$time,
                      levels = c(1,2,3,4,5,6),
                      labels = c("29.9.2015", "23.10.2015", "15.11.2015", "22.4.2016", "12.9.2016", "xx.10.2017"))


```


```{r, message=FALSE, warning=FALSE}
ggpairs(
  Kuitub, c(7,9,13,15),
  lower = list(combo = wrap("facethist", binwidth = 0.5)))
```

Nitraattitypen ja parametrien väliset erot

```{r, message=FALSE, warning=FALSE}
p <- ggplot(Kuitub, aes(kas, nitN))
p + geom_boxplot(aes(fill = lan)) +theme_tufte()+geom_rangeframe(color="black") +
  ggtitle("All measurement average nitrate NO3- (mg/kg) with standard deviation")
```

Kaikkien käsittelyiden lannoituskäsittelyiden keskimääräiset erot

```{r, message=FALSE, warning=FALSE}
Kuitub2 = Kuitub[ which(Kuitub$time=="xx.10.2017"), ]
p <- ggplot(Kuitub2, aes(kas, nitN))
p + geom_boxplot(aes(fill = lan)) +theme_tufte()+geom_rangeframe(color="black") +
  ggtitle("xx.10.2017 Nitrate NO3- (mg/kg) means with standard deviation")

```

Viimeisen mittausajankohdan käsittelyjen erot - huomaa erot paljon pienempiä


```{r, message=FALSE, warning=FALSE}
p <- ggplot(Kuitub, aes(lan, LisC))
p + geom_point(aes(fill = lan)) +theme_tufte()+geom_rangeframe(color="black") +
  ggtitle("Lisätty hiili lannoitteissa (kg/ha)")
```

**TOC/N**
In table:

Käsittelyt    | S1  |S2 | S3 | S4 | Min |  C 
--------    |----|----|----|----|-----|----
Käsittelyt                | S1 | S2  | S3 | S4  | Min | C 
Kokonaistyppi             | 298 | 231 | 230 | 355 | 0 |  0      
Lisätty liukoinen typpi   | 17 | 15 |  28 | 28  | 0 |  0
TOC/N                     | 66  | 77  | 55  | 61  | <1  | 0
TOC/Liukoinen N           | 1145 | 1159 | 444 | 773 | <1  |  0

Vaikka lannoitteet sisältävät eri määrän liukoista typpeä ovat niiden hiilityppisuhteet (TOC/N) verrattain samankaltaisia.

```{r, message=FALSE, warning=FALSE}
agg = ddply(Kuitub, .(kas, lan, time), function(x){
  c(mean=mean(x$nitN), sd = sd(x$nitN))
})

#agg #contains all nitN treatment means and standard deviations
#write.table(agg, "nitNmeans6times.txt", sep="\t")

agg$lower = agg$mean + agg$sd
agg$upper = agg$mean - agg$sd

#agg #contains mean, sd, upr, and lwr confidence levels

pd <- position_dodge(width = 0.2) # move them .2 to the left and right

#FIRST RUN ONE OF THESE

#Plot2x4
gbase  = ggplot(agg, aes(y=mean, colour=lan)) +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=.3, position=pd) +
  geom_point(position=pd) + facet_wrap(~ kas, ncol=2)
gline = gbase + geom_line(position=pd) + ylim(0, 25)

agg$num_time = NA
agg$num_time = as.numeric(agg$time)

#for all graphs
gline = gline %+% agg

gline_final = gline + aes(x=num_time) + scale_x_continuous(breaks= 1:length(agg$time), labels = agg$time) +
  theme(axis.text.x = element_text(angle=60, hjust=1)) +
  labs(x = "Time", y = "Nitrate NO3- (mg/kg) means with confidence intervals")

#Final graph
print(gline_final+theme_tufte()+geom_rangeframe(color="black"))

```

Kaikki käsittelyt graafisesti esitettynä. Erot ovat aluksi suurempia ja pienevät kohti kokeen loppua. 12.9.2016 erot ovat toisinpäin. Todennäköisesti typpi on alkanut vapautua orgaanisesta aineesta.

## Analyysit

```{r, message=FALSE, warning=FALSE}
# DATA...
Kuitub1 = Kuitub[ which(Kuitub$time=='15.11.2015' & Kuitub$kas == "R" ), ]
```

```{r, message=FALSE, warning=FALSE}
summary_data <- ddply(Kuitub1,
                      c("lan"),
                      summarise,
                      N = length(nitN),
                      avg_value = mean(nitN),
                      sd = sd(nitN),
                      se = sd/sqrt(N)
)

# check out the summary stats 
summary_data
```

Käsittelyjen keskiarvot, keskihajonnat sekä keskivirheet.

```{r, message=FALSE, warning=FALSE}
# Plotting
raw_plot <- ggplot(Kuitub1, aes(x=lan,y=nitN))
raw_plot <- raw_plot + geom_point(position = position_jitter(w=0.2),alpha=0.4, colour="grey58")
raw_plot <- raw_plot + xlab("Treatment")
raw_plot <- raw_plot + ylab("Value")
raw_plot <- raw_plot #+ ylim(0,15)
raw_plot <- raw_plot + theme_tufte(base_size=18)+geom_rangeframe(color="black")


# Kuitub1$lan <- factor(Kuitub1$lan, levels=c("S1", "S2", "S3", "S4", "Min", "C"))
# To change the order, not needed now.

# we'll also do this for the summary data and use it later on 
# summary_data$treatment <- factor(summary_data$treatment, levels=("S1", "S2", "S3", "S4", "Min", "C"))

#Make labels
summary_data$labels <- c("a","a","a","a","b","c")


combined_plot <- raw_plot + geom_point(data=summary_data, aes(x=lan,y=avg_value), colour="black", size=3)
combined_plot <- combined_plot + geom_errorbar(data = summary_data,
                                               aes(x = lan, y = avg_value,
                                                   ymax = avg_value + sd,ymin = avg_value - sd), 
                                               width=0.1, colour="black")
combined_plot <- combined_plot + annotate("text",x="S1",y=0,label="n=4")
combined_plot <- combined_plot + annotate("text",x="S2",y=0,label="n=4")
combined_plot <- combined_plot + annotate("text",x="S3",y=0,label="n=4")
combined_plot <- combined_plot + annotate("text",x="S4",y=0,label="n=4")
combined_plot <- combined_plot + annotate("text",x="Min",y=0,label="n=4")
combined_plot <- combined_plot + annotate("text",x="C",y=0,label="n=4")
combined_plot <- combined_plot + geom_text(data=summary_data,aes(x=lan,y=avg_value+se+3,label=labels))
combined_plot + ggtitle("R + nitrate levels @ 15.11.2015")
```

Käsittelyt S1-S4 ovat samoja, mineraalilannoite ja kontrolli eroavat toisistaan

```{r, message=FALSE, warning=FALSE}
fit <- aov(nitN ~ factor(lan) + factor(ker),
           data = Kuitub1)
TukeyHSD(fit, "factor(lan)", conf.level = 0.95)

```

```{r, message=FALSE, warning=FALSE}
```

```{r, message=FALSE, warning=FALSE}
```




